Mathematical Models of Delta-Notch Signalling

### 2025-03-15 paper/paper.tex
Added SDE section to paper.tex

>>>>

To capture the inherent randomness in cellular processes while maintaining computational efficiency, we implement a stochastic differential equation (SDE) model based on the chemical Langevin equation formalism. This approach represents an intermediate level of detail between the deterministic ODE model and the fully discrete Gillespie algorithm.

Starting from the deterministic ODEs described in Section 3.3, we incorporate stochastic fluctuations by adding noise terms proportional to the square root of the reaction propensities:
$$
\begin{aligned}
  dN &= \left( \frac{n_{m}I^2}{n_{0}^2 + I^2} - k_{T}ND_{ext} - \gamma N \right) dt + \sigma \cdot N \cdot dW_{N} \\ \\[5pt]
  dD &= \left( \frac{d_{m}d_{0}^2}{d_{0}^2 + I^2} - k_{T}DN_{ext} - \gamma D \right)  dt + \sigma \cdot D \cdot dW_{D}  \\[5pt]
  dI &= \left( k_{T}ND_{ext} - \gamma_{I}I \right) dt + \sigma \cdot I \cdot dW_{I}
\end{aligned}
$$
where $dW_N$, $dW_D$, and $dW_I$ are independent Wiener processes representing white noise. The structure of the noise terms follows from the assumption that noise in each reaction channel is proportional to the current concentration of those molecules. \par

For numerical implementation, we employ the Euler-Maruyama scheme with time step $dt$. The deterministic step is first calculated, and then the Wiener increments are sampled from a normal distribution with mean 0 and standard deviation $\sqrt{dt}$. Each increment is scaled by the its respective variable and added to the results of the deterministic step.
